\section{Implementation and Experiment Setup}
\label{sect:implementation-and-experiment-setup}

Besides this written report, we have implemented the framework illustrated in 
\crefrange{chap:visualizing-static-input-graphs}{chap:visualizing-dynamic-input-graphs}, along with the test case generation discussed in \cref{sect:test-case-generation}, as a Swift package.
We also provide a graphical user interface to explore the individual steps of the pipeline interactively.

To perform the experimental evaluation, we include a command-line tool that efficiently generates test instances, runs them through the pipeline, and assesses their quality.
Each instance is generated randomly based on the parameters $n$, $\alpha$, and $\beta$ discussed previously; the bounding box $\mathcal{A}$ and the weight distribution $\pmf$ are kept constant.
We vary $n$, $\alpha$, and $\beta$ to get insights into their effect on the generated maps' qualities.
Whenever we run the force-directed optimization algorithm, we perform $10c$ iterations.
Here, $c$ is the current number of clusters in the cluster graph or the number of regions in the polygonal dual.

The implementation is open source and can be found on GitHub:
%
\begin{center}
  \url{https://github.com/jenox/master-thesis/}
\end{center}

The results presented here are based on 100 randomly generated identifiers that are included in the repository.
We use these identifiers to seed the pseudorandom number generator used by our framework, thereby making the results presented here reproducible on any machine running macOS 10.15 or later.
