% https://www.scribbr.com/dissertation/write-conclusion/
\chapter{Conclusion}
\label{chap:conclusion}

% Summarize and reflect on the research
In this thesis, we explored how cluster information in graphs can be visualized naturally as countries on a map in a dynamic setting where the graph changes over time.
We proposed a framework that produces maps whose regions' sizes are approximately proportional to the respective clusters' sizes.
The framework can also react to dynamic changes of the input graph in a way that preserves the viewer's mental model of the map.

% Show what new knowledge you have contributed
We can make these guarantees by working on an intermediate representation that encompasses all combinatorial properties of the eventual map, breaking changes down into small, incremental changes, and using a force-directed algorithm to draw the eventual map.
This way, the structure of our algorithmic pipeline alone guarantees that changes to the map are minor and comprehensible, thereby preserving the viewer's mental map as the underlying data changes over time.

Besides the preservation of the mental map, related work struggles with keeping clusters as continuous regions in the generated maps \cite{gansner2009gmap} \cite{mashima2011visualizing}.
The structure of our algorithmic pipeline addresses and solves this problem, too.
Considering the maps we produce are contact representations of the intermediate cluster graph, clusters cannot possibly be fragmented across different regions.

The framework presented here can be applied to a variety of real-world applications, allowing us to visualize clusters in the underlying data and how they change over time, thereby enabling viewers to detect trends in the data easily.

% Clearly state the answer to the main research question

% Make recommendations for future work on the topic
\input{Sources/06-01-Future-Work.tex}
